### 2012-09-10 draft.tex
added Introduction

\documentclass[times, 10pt, twocolumn]{article}
\usepackage{latex8}
\usepackage[dvips]{graphicx}
\usepackage[fleqn]{amsmath}
\usepackage{amsthm}
\usepackage{txfonts}
\usepackage{courier}
\usepackage{subfigure}
\usepackage{comment}
\usepackage{url}
\usepackage[square,numbers]{natbib}

\newcommand{\hd}{\emph{H+D}}
\newcommand{\hp}{\emph{Hpin}}
\newcommand{\dm}{\emph{Dmap}}

%-----------------------------------------
% need for camera-ready
%\pagestyle{empty}

%----------------------------------------

\title{Accelerating Traffic Simulation and Its Real-Time Challenges}

\author {
Manato Hirabayashi, Shinpei Kato, Masato Edahiro\\
\textit{Department of Information Engineering}\\
\textit{Nagoya University}\\
\and
Yuki Sugiyama\\
\textit{Department of Physics}\\
\textit{Nagoya University}\\
}

\begin{document}

\maketitle

%-----------------------------------------
% need for camera-ready
%\thispagestyle{empty}

\begin{abstract}
\end{abstract}

\section{Introduction}

Transportation systems underlie our industry and life as part of
societal infrastructure.
Economy depends highly on traffic flow.
Once a traffic jam occurs, for example, it could cause significant
economic damage through degradation of transport efficiency, energy
consumption, and environmental poisoning.
According to the Japanese government report a few years ago, economic
losses due to traffic congestion reach one hundred billion dollars per
year in Japan.
Albeit a major issue of transportation systems, the mechanism of traffic
jam is not well-explained in the literature.
What is particularly important is ``phantom jam'', which often happens
on freeways without any accident.
This phantom jam should be avoided by science, given that it is
a physical phenomenon but not one caused by human errors.
However, we are not aware of scientific and engineering solutions to
such real-world problems.

In physics, traffic flow is described by mathematical
models~\cite{Bando1995, Kerner1993, Nagel1992}.
While their mathematical expressions are not identical, they are all
compute-intensive procedures.
For example, traffic simulation based on the Optimal Velocity (OV)
model~\cite{Bando1995} computes locations and velocities of agents every
sampling period, solving OV equations. 
Specifically, the location $x_n$ of the $n$th agent is described by the
following equation, where $\Delta x_n = x_{n+1} - x_n$ is a distance to
a preceding agent, $a$ is a sensitivity, and $V()$ is an optimal
velocity function:
\begin{eqnarray}
 \label{eqn:ov}
 \frac{d^2 x_n}{d t^2} = a \left\{V(\Delta x_n) - \frac{d x_n}{d t}\right\}.
\end{eqnarray}

Applying a large number of agents to simulation using the above formula,
it is apparent that computational workload increases exponentially.
In fact, the above formula considers only one dimension, which restricts
applications of simulation to freeway traffic flows, powder flows,
molecular motors, and so on.
Making it multi-dimensional further increases computational workload,
while allowing more complicated simulations, such as traffic networks,
internet packet flows, evacuation routes, and herd formations of
animals, to use similar optimal velocity models.
Given a scale of million agents in the real world, traffic
simulation should be supported by powerful computer systems.

Deployment of traffic simulation may require real-time feedback from the
real world.
This turns out to be a cyber-physical systems (CPS) problem.
Real-time traffic simulation is another challenging issue, where
the rate and the preciseness of simulation must be traded to meet the
requirement of a given scenario.
For instance, evacuation may want very high rate simulation, sacrificing
the preciseness of simulation to some extent.
Unfortunately, none of those challenges has been explored yet, largely
due to a lack of multidisplinary research collaborations between physics
and computer science.

This paper explores how to accelerate traffic simulation using the
state-of-the-art parallel computing technology.
In particular, we use the graphics processing unit (GPU), which
integrates hundreds of cores on a chip.
Recent GPUs are becoming more and more suitable for general-purpose
data-parallel applications.
The traffic simulation program used in this paper applies
Equation~\eqref{eqn:ov} to a large number of agents.
The resulting workload is highly data-parallel and compute-intensive,
which can be nicely offloaded on to the GPU.
We also identify the bottleneck in accelerating our traffic simulation
program using the GPU, and provide some insight into its solution.

\section{Assumption}
\label{sec:assumption}

\begin{enumerate}
 \item Initalize the time $t$, and set the initial values of $x_n(t)$
       (also $y_n(t)$ and $z_n(t)$, if necessary for multidimensional
       versions).
 \item Increase the time $t$ by the sampling period $\Delta t$.
 \item Compute the location $x_n(t)$ (also $y_n(t)$ and $z_n(t)$, if
       necessary for multidimensional versions), and the velocity
       $v_x(t)$ at time $t$ for each agent $A_n$, using the OV model.
 \item Go back to Step 2, if the simulation time is expired.
 \item Exit the program.
\end{enumerate}

Note that Step 3 explots a lot of loop procedures to derive the
locations and velocities of agents.
This part can be accelerated by the GPU.

\section{GPU Implementation}
\label{sec:gpu_implementations}

GPU performance is dominated by the program design.
GPU-accelerated programs are typically divided into two pieces of code.
The CPU code plays a role of a master thread that controls the program
flow.
The GPU code, on the other hand, spawns a bunch of worker threads to
execute compute-intensive parts of the program in parallel, thus
accelerating the overall program.
What is often argued to optimize performance is how to parallelize the
program into threads.

In order to optimize GPU performance, however, we need to make a
difficult decision about when to offload on to the GPU.

\begin{figure}[t]
\centering
\includegraphics[width=0.4\textwidth]{eps/sample-in-cpu.eps}
\caption{Sampling by the CPU.}
\label{fig:sample-in-cpu}
\end{figure}

\begin{figure}[t]
\centering
\includegraphics[width=0.4\textwidth]{eps/sample-in-gpu.eps}
\caption{Sampling by the GPU.}
\label{fig:sample-in-cpu}
\end{figure}

\section{Evaluation}
\label{sec:evaluation}

\begin{figure}[t]
\centering
\includegraphics[width=0.5\textwidth]{eps/eval_accel.eps}
\caption{Acceleration of traffic simulation.}
\label{fig:eval_accel}
\end{figure}

\begin{figure}[t]
\centering
\includegraphics[width=0.5\textwidth]{eps/eval_nosync.eps}
\caption{Impact of synchronization.}
\label{fig:eval_nosync}
\end{figure}

\begin{figure}[t]
\centering
\includegraphics[width=0.5\textwidth]{eps/eval_error.eps}
\caption{Error in location and velocity due to imprecise synchronization.}
\label{fig:eval_error}
\end{figure}

\section{Conclusion}
\label{sec:conclusion}

Gdev~\cite{Kato2012}.

\bibliographystyle{plain}
{\footnotesize
\bibliography{references}
}

\end{document}

### 2012-09-11 draft.tex
added GPU Implementation

@@
 Unfortunately, none of those challenges has been explored yet, largely
-due to a lack of multidisplinary research collaborations between physics
+due to a lack of multidisciplinary research collaborations between physics
 and computer science.
@@
 
+grid, block, thread.
+
+A unit of functions that is launched on the GPU is called a \textit{kernel}.
+
+\section{Traffic Simulation}
+\label{sec:traffic_simulation}
+
 \begin{enumerate}
- \item Initalize the time $t$, and set the initial values of $x_n(t)$
+ \item Initialize the time $t$, and set the initial values of $x_n(t)$
        (also $y_n(t)$ and $z_n(t)$, if necessary for multidimensional
@@
 
-Note that Step 3 explots a lot of loop procedures to derive the
+Note that Step 3 exploits a lot of loop procedures to derive the
 locations and velocities of agents.
@@
 accelerating the overall program.
-What is often argued to optimize performance is how to parallelize the
-program into threads.
+What is often argued in performance optimization is how to parallelize
+the compute-intensive parts into threads.
+This is actually the well-studied problem in the literature.
+What is not really understood yet is when to offload the program on to
+the GPU.
+This paper explores two schemes that use the GPU at different timings to
+see how GPU performance is affected.
 
-In order to optimize GPU performance, however, we need to make a
-difficult decision about when to offload on to the GPU.
+\subsection{Sample-in-CPU Scheme}
+\label{sec:sample-in-cpu}
+
+We first implement such a scheme that uses the GPU \textit{only if
+necessary}.
+In other words, the program is offloaded on to the GPU, only when
+computations of locations and velocities of agents are accelerated by
+parallelization.
+The CPU bridges across sampling periods to manage simulation.
+In this scheme, the control flow of simulation is always returned to the
+CPU at the end of each sampling period.
+Such a synchronized approach makes the programmer easy to
+obtain intermediate results of simulation, whereas the overhead imposed
+on moving back and forth between the CPU and the GPU must be
+compromised.
 
@@
 \includegraphics[width=0.4\textwidth]{eps/sample-in-cpu.eps}
-\caption{Sampling by the CPU.}
+\caption{Block diagram of the Sample-in-CPU scheme.}
 \label{fig:sample-in-cpu}
 \end{figure}
+
+Figure~\ref{fig:sample-in-cpu} shows a brief overview of the
+Sample-in-CPU scheme.
+This scheme also has several alternatives depending on how many kernels
+need to be launched on the GPU in each period. 
+Suppose that we want to offload $n$ pieces of \textit{for} loops on to
+the GPU.
+We may return to the CPU $n$ times in totall, that is, return every time
+one \textit{for} loop breaks, or otherwise we may just return once when
+all the $n$ \textit{for} loops end.
+This is a design decision, and is also dependent on the GPU architecture
+and the parallelization structure.
+In general, parallelized threads need to be synchronized when moving
+across basic blocks.
+Specifically, when moving to the next \textit{for} loop executed in
+parallel, all the threads relevant to this parallel computing procedure
+must synchronize with each other.
+However, the maximum number of threads that can be synchronized on the
+GPU is often limited.
+As of 2012, for example, NVIDIA's GPU architectures limit the number of
+such threads to 1024 or less.
+Our implementation therefore forces the program to return to the CPU
+every time one \textit{for} loop breaks.
+Note that this is not a conceptual limitation of GPU computing, but is a
+current limitation of hardware.
+We believe that this limitation would be removed or mitigated in
+next-generation GPU architectures.
+
+\subsection{Sample-in-GPU Scheme}
+\label{sec:sample-in-gpu}
 
@@
 \includegraphics[width=0.4\textwidth]{eps/sample-in-gpu.eps}
-\caption{Sampling by the GPU.}
-\label{fig:sample-in-cpu}
+\caption{Block diagram of the Sample-in-GPU scheme.}
+\label{fig:sample-in-gpu}
 \end{figure}
+
+We next implement such a scheme that uses the GPU \textit{all the time},
+even to control simulation.
+There is one big kernel running on the GPU, which is launched only once
+at the beginning.
+After offloading the program on to the GPU, the CPU is going to wait for
+the completion of simulation.
+This scheme is almost optimized in performance, since there is little
+overhead in communication between the CPU and the GPU.
+The downside of this scheme is that the CPU and the GPU are not
+synchronized.
+The progress of simulation is not visible from the CPU, unless the
+program implements a specific interface to allow the CPU to access
+intermediate results of simulation running on the GPU.
+In our implementation, we consider providing a framework that the CPU
+downloads data from the GPU asynchronously without awareness of
+simulation, when the user requests intermediate results.
+This asynchronous data access may affect the performance of simulation,
+though in reality this access would not happen more than once in a
+sampling period.
+
+Figure~\ref{fig:sample-in-gpu} shows a block diagram of the
+Sample-in-GPU scheme.
+The procedure is very simple.
+The most portion of code of simulation is executed on the GPU.
+Given that the single-thread performance of recent GPUs is getting more
+reliable, this approach is pretty reasonable.
+As mentioned in Section~\ref{sec:sample-in-cpu}, however, the current
+limitation of the GPU architecture prevents us from synchronizing among
+blocks at a scale of thousands threads.
+Therefore, this scheme is speculative in a sense that it does not work today
+but may appear in the future.
+
+This paper provides some degree of insights into how this scheme is
+effective.
+We implement this scheme with the current GPU architecture under the
+assumption that global synchronization among blocks works.
+There is also another possible approach for this scheme that we limit
+the number of threads to what is supported by the GPU architecture.
+This alternative implementation, however, is left open for future work.
 
@@
 \label{sec:evaluation}
+
+We evaluate the performance of our GPU-accelerated traffic simulation
+programs, using an NVIDIA GeForce GTX 560 Ti graphics card.
+This is a middle-end graphics card integrating 394 compute cores on a
+chip.
 
@@
 
+\subsection{Experimental Results}
+\label{sec:experimental_results}
+
+\subsection{Discussion}
+\label{sec:discussion}
+
+Imprecise Computation.
+Gdev~\cite{Kato2012}.
+
 \section{Conclusion}
 \label{sec:conclusion}
-
-Gdev~\cite{Kato2012}.
 

### 2012-09-12 draft.tex
added Evaluation

@@
 \begin{abstract}
+PLACEHOLDER.
 \end{abstract}
@@
 
-\section{Assumption}
+The rest of this paper is organized as follows.
+Section~\ref{sec:assumption} describes the assumption and terminology
+used this paper.
+Section~\ref{sec:traffic_simulation} briefly explains our traffic
+simulation program.
+Section~\ref{sec:gpu_implementations} presents our schemes of GPU
+implementations of traffic simulation.
+Section~\ref{sec:evaluation} evaluates the benefit of our GPU
+implementations, and provides an insight into future work.
+Section~\ref{sec:conclusion} concludes this paper.
+
+\section{Assumption and Terminology}
 \label{sec:assumption}
 
-grid, block, thread.
+We assume the Compute Unified Device Architecture (CUDA) for GPU
+programming.
+In CUDA, a unit of pieces of code that is launched on the GPU is called
+a \textit{kernel}.
+The kernel is typically composed of multiple parallelized
+\textit{threads} that accelerate computations.
+A unit of threads is called a \textit{block} where most computing
+resources are shared.
+A collection of blocks for the corresponding kernel is called a
+\textit{grid}. 
+The maximum number of threads that an individual block can contain is
+defined by the GPU architecture.
 
-A unit of functions that is launched on the GPU is called a \textit{kernel}.
+GPU applications use a set of the API supported by CUDA.
+We typically take the following steps to use the GPU: (i) allocate space
+to device memory, (ii) copy data to the allocated device memory space,
+(iii) launch the program on the GPU, (iv) copy resultant data back to
+host memory, and (v) free the allocated device memory space.
 
@@
 \label{sec:traffic_simulation}
+
+PLACEHOLDER.
 
@@
 As of 2012, for example, NVIDIA's GPU architectures limit the number of
-such threads to 1024 or less.
+such threads to 1024 or less~\cite{NVIDIA_Kepler}.
 Our implementation therefore forces the program to return to the CPU
@@
 programs, using an NVIDIA GeForce GTX 560 Ti graphics card.
-This is a middle-end graphics card integrating 394 compute cores on a
-chip.
+This is a middle-end graphics card based on the Fermi
+architecture~\cite{NVIDIA_Fermi}, integrating 394 compute cores on a chip.
+The programs used in this evaluation are written in CUDA, and are
+compiled by the NVIDIA CUDA Compiler (NVCC) v4.2 suite~\cite{CUDA42}.
+
+We focus on the one-dimensional OV model~\cite{Bando1995}, which can be
+applied to freeways simulation.
+The original program~\cite{OV_Program} of this simulation is written in
+C, without any parallelization.
+We use the same simulation parameters as the original default setup,
+while increasing the number of agents (cars) to see how simulation
+performance varies. 
+The evaluation is conducted by comparing our GPU implementations with
+the original CPU implementation.
+
+\subsection{Performance Benefit}
 
@@
 \includegraphics[width=0.5\textwidth]{eps/eval_accel.eps}
-\caption{Acceleration of traffic simulation.}
-\label{fig:eval_accel}
+\caption{Performance benefit of the GPU.}
+\label{fig:eval_benefit}
 \end{figure}
+
+Figure~\ref{fig:eval_benefit} demonstrates the performance benefit of
+the GPU under the current hardware limitation.
+The Sample-in-CPU scheme herein exploits synchronization between the CPU
+and the GPU after every \textit{for} loop to overcome the limitation that
+CUDA threads cannot synchronize beyond blocks.
+Despite of this restricted implementation, our GPU implementation
+outperforms the original CPU implementation by about five times in
+simulation time, when the number of agents exceeds a scale of 1000.
+This evinces an affinity of the GPU and parallel computing for traffic
+simulation.
+It is also interesting to see that the CPU implementation has a better
+performance for a small number of agents.
+Thus, the overhead of communication between the CPU and the GPU has
+non-trivial impact if the achievable parallelism is not sufficient.
+Another notable observation is that the diffence in performance of the
+CPU and the GPU implementations is saturated when the number of agents
+reaches 1000.
+This saturation implies the hardware potential of the NVIDIA GeForce GTX
+560 Ti graphics card used in this evaluation.
+Using high-end graphics cards, the difference in performance may not be
+saturated at the same scale.
+
+\subsection{Impact of Hardware Limitation}
 
@@
 \includegraphics[width=0.5\textwidth]{eps/eval_nosync.eps}
-\caption{Impact of synchronization.}
+\caption{Impact of the hardware limitation.}
 \label{fig:eval_nosync}
 \end{figure}
+
+Figure~\ref{fig:eval_nosync} demonstrates the impact of the current
+hardware limitation that CUDA threads cannot synchronize beyond blocks.
+In this experiment, we observe ``what would happen if this hardware
+limitation is removed?'', compromising the preciseness of simulation
+results.
+Namely, we use the current CUDA synchronization function to synchronize
+threads running in different blocks, even though it does not work as
+expected.
+All the GPU implementations of the plotted schemes use this assumption.
+Therefore, the Sample-in-CPU scheme is now different from what we have
+evaluated in Figure~\ref{fig:eval_benefit}, since there is no need
+anymore to return to the CPU every time one \textit{for} loop breaks.
+We instead put a synchronization function between \textit{for} loops on
+the GPU.
+The Sample-in-CPU-Sync scheme is an alternative version of the
+Sample-in-CPU scheme in that we download intermediate results from the
+GPU to the CPU at the end of each sampling period.
+The Sample-in-GPU scheme is what is presented in
+Section~\ref{sec:sample-in-gpu}.
+
+It is very notable that the Sample-in-GPU scheme improves the simulation
+time by a factor of 1000 as compared to the original CPU implementation.
+This result encourage suport for thread synchronization among blocks in
+CUDA programming.
+There are also several interesting observations obtained in this
+experiment.
+Comparing the Sample-in-CPU and the Sample-in-CPU-Sync schemes, it turns
+out that the cost of downloading intermediate results from the GPU to
+the CPU is becoming trivial as the scale of simulation increases.
+Hence, a common argument of ``host-device data communication in GPU
+programming is expensive'' does not apply to large-scale traffic
+simulation.
+Another issue of concern often discussed in GPU programming is that
+a single-thread performance of GPUs is weak as compared to that of CPUs.
+This is true in general, but our experimental result shows that the
+performance loss caused by pushing most pieces of code into the GPU in
+the Sample-in-GPU scheme is not signficant even for a small number of
+agents.
+This observation leads to some conclusion that the Sample-in-GPU scheme
+would be the best choice for any scale of traffic simulation.
+
+\subsection{Practical Concern}
 
@@
 
-\subsection{Experimental Results}
-\label{sec:experimental_results}
+We finally examine the impact of misbehavior of thread synchronization
+in our GPU implementations.
+The simulation results produced by our GPU implementations are not
+precise except for those of the Sample-in-CPU scheme presented in
+Figure~\ref{fig:eval_benefit}, due to our optimistic use of thread
+synchronization on the GPU.
+We need to wait for new GPU architectures removing the limitation of
+thread synchronization in order to make them precise.
+Practically speaking, however, imprecise results of simulation are still
+valuable and meaningful as far as they are limited to an acceptable
+range of errors for traffic flow.
+
+Figure~\ref{fig:eval_error} demonstrates the error percentage of
+location and velocity between the simulation results produced by the
+original CPU and the imprecise Sample-in-GPU implementations.
+Since this is one-dimensional OV simulation, there are only forward or
+backward in location, and high or low in velocity.
+Under this constraint, the error in location is very trivial.
+This is because the characteristic of traffic flow is well captured,
+even though the simulated locations and velocities have some error in
+their individual values.
+A degree of the error in velocity is slightly high as compared to that
+in location.
+This means that velocity is more sensitive to imprecise results than
+location.
+However, the maximum error in velocity observed in our experiment is at
+most 5\%.
+Suppose that the average driving speed is 50mph.
+The simulated speed in our imprecise implementation would still be
+within a range of 47.5 to 52.5mph.
+Intuitively, this scale of error is acceptable in practice.
+We plan to further investigate if the same argument can be applied to a
+more variety of parameter setups and multidimensional OV simulation
+programs.
 
@@
 
-Imprecise Computation.
+PLACEHOLDER.
+
+Imprecise Computation~\cite{Lin1987}.
 Gdev~\cite{Kato2012}.
@@
 \label{sec:conclusion}
+
+PLACEHOLDER.
 

### 2012-09-12 draft.tex
completed the first draft

@@
 \author {
-Manato Hirabayashi, Shinpei Kato, Masato Edahiro\\
+Manato Hirabayashi, Shinpei Kato and Masato Edahiro\\
 \textit{Department of Information Engineering}\\
@@
 \begin{abstract}
-PLACEHOLDER.
+ Traffic simulation is a growing domain of computational physics.
+ There are many applications in our industry and life that benefit from
+ traffic simulation.
+ A core challenge of this research domain, however, is posed due to
+ an unbounded scale of computation.
+ In this paper, we explore an advantage of using the graphics processing
+ unit (GPU) to support its computation.
+ Two schemes are presented to study how to maximize GPU
+ performance in this context.
+ Our evaluation using a commodity graphics card shows that our GPU
+ implementation improves simulation time by about five times over the
+ original CPU implementation.
+ We also demonstrate by experiments that additional orders-of-magnitude
+ improvements could be achieved by overcoming the current hardware
+ limitation.
 \end{abstract}
@@
 a \textit{kernel}.
-The kernel is typically composed of multiple parallelized
-\textit{threads} that accelerate computations.
-A unit of threads is called a \textit{block} where most computing
-resources are shared.
-A collection of blocks for the corresponding kernel is called a
-\textit{grid}. 
-The maximum number of threads that an individual block can contain is
-defined by the GPU architecture.
+The kernel is typically composed of multiple \textit{threads} that
+executes the code in parallel.
+A unit of threads that are co-scheduled by hardware is called a
+\textit{block}, while a collection of blocks for the corresponding
+kernel is called a \textit{grid}.  
+The maximum number of threads that can be contained by an individual
+block is defined by the GPU architecture.
 
-GPU applications use a set of the API supported by CUDA.
-We typically take the following steps to use the GPU: (i) allocate space
+CUDA programs use a set of the application programming interface (API)
+functions to control the GPU.
+We usually take the following steps to use the GPU: (i) allocate space
 to device memory, (ii) copy data to the allocated device memory space,
@@
 
+Our computing platform is composed of a single CPU and GPU.
+The traffic simulation program is a single process from the CPU point of
+view.
+Hence, kernels are launched from the CPU sequentially, while each kernel
+is executed by parallelized threads on the GPU.
+
 \section{Traffic Simulation}
@@
 
-PLACEHOLDER.
+The motivation to study traffic flow is to understand the property of
+traffic flow, mainly why traffic congestion occurs.
+The study of traffic flow has a long history, and many engineering and
+physical models exist.
+This paper is particularly focused on the OV model~\cite{Bando1995}.
+In this model, a driver is supposed to maintain an optimal velocity that
+depends on the driver's headway.
+Its motion is given by Equation~\eqref{eqn:ov}.
+The optimal velocity function introduced in this equation is represented
+by the relationship between the driver's headway and velocity.
+
+The OV model has a simple form with a single parameter of sensitivity
+$a$.
+This simplicity allows us to understand the mechanisms of phenomena, and
+also predict the results of experiments.
+The disadvantage of this model, on the other hand, is a little allowance
+for control systems to turn up the parameter.
+There is indeed a need for collaborations between the research domains
+of engineering and physics.
+To begin with, however, we restrict our attention to the OV model.
+More details are found in~\cite{Bando1995}.
+
+The traffic simulation program based on the OV model is straightforward,
+as the model itself has a simple form:
 
@@
 
-Note that Step 3 exploits a lot of loop procedures to derive the
+The computation in Step 3 exploits a lot of loop procedures to derive
 locations and velocities of agents.
-This part can be accelerated by the GPU.
+This part is the most compute-intensive, and thus can be accelerated by
+the GPU.
 
@@
 the GPU.
-We may return to the CPU $n$ times in totall, that is, return every time
+We may return to the CPU $n$ times in total, that is, return every time
 one \textit{for} loop breaks, or otherwise we may just return once when
@@
 non-trivial impact if the achievable parallelism is not sufficient.
-Another notable observation is that the diffence in performance of the
+Another notable observation is that the difference in performance of the
 CPU and the GPU implementations is saturated when the number of agents
@@
 time by a factor of 1000 as compared to the original CPU implementation.
-This result encourage suport for thread synchronization among blocks in
+This result encourage support for thread synchronization among blocks in
 CUDA programming.
@@
 performance loss caused by pushing most pieces of code into the GPU in
-the Sample-in-GPU scheme is not signficant even for a small number of
+the Sample-in-GPU scheme is not significant even for a small number of
 agents.
@@
 
-PLACEHOLDER.
+According to the experimental results, it is highly desired to use the
+Sample-in-GPU scheme, if available.
+It provides an orders-of-magnitude improvement in simulation time.
+The downside of this scheme at the moment is a lack of functionality in
+thread synchronization due to the current hardware limitation.
+Simulation results are never precise unless thread synchronization
+operates correctly.
+Although we have demonstrated that the error in simulation results
+caused by a lack of thread synchronization is acceptable in our
+experimental setup, the range of error is never bounded as depicted in
+Figure~\ref{fig:eval_error}.
+Hence, the availability of this scheme depends highly on practical
+setups.
 
-Imprecise Computation~\cite{Lin1987}.
-Gdev~\cite{Kato2012}.
+We consider that the problem of this scheme may be relevant to the
+Imprecise Computation model~\cite{Lin1987}.
+The Sample-in-CPU scheme is precise-but-slow, while the Sample-in-GPU
+scheme is imprecise-but-fast.
+If we are allowed to employ multiple GPUs in the system, we may run the
+two types of simulation programs on different GPUs concurrently.
+At some point of time $t$, the simulation result produced by the
+Sample-in-GPU scheme represents a scenario way ahead of that produced by
+the Sample-in-CPU scheme; the question is how precise it is.
+Figure~\ref{fig:eval_error} demonstrates that the error in the
+simulation result produced by the Sample-in-GPU scheme is getting larger
+as time goes by.
+If we can find such a time duration that causes the value of error to
+fall out the acceptable range for practical use, we define this time
+duration to be $\Delta t$.
+Suppose that the two simulation programs based on the Sample-in-GPU and
+the Sample-in-CPU schemes, respectively, are running concurrently.
+At every interval of time $\Delta t$, we reset the simulation result
+produced by the Sample-in-GPU scheme, and input the latest precise
+simulation result produced by the Sample-in-CPU scheme during the same
+interval.
+By this means, the value of error in the simulation result produced by
+the Sample-in-GPU scheme never falls out the acceptable range, while we
+still benefit from its simulation speed within an internal of time
+$\Delta t$.
+The value of $\Delta t$ can be determined adaptively based on the
+Imprecise Computation model.
+
+So far we have studied a single process of traffic flow.
+In reality, however, there are multiple processes of traffic flow.
+Real-time feedback from the real road conditions may also be
+additionally required to deploy traffic simulation.
+All these pieces of simulation must meet the quality-of-service (QoS)
+requirement for users.
+Now, this turns into a real-time systems problem.
+We have developed a lot of solutions to use the GPU in real-time
+systems~\cite{Kato2011_3, Kato2011_2, Kato2011_1, Kato2012}.
+We believe that some of these work could contribute to GPU-accelerated
+real-time traffic simulation.
 
@@
 
-PLACEHOLDER.
+In this paper, we have presented GPU-accelerated traffic simulation
+based on the OV model.
+The current hardware limitation of the GPU architecture caps
+improvements in performance to a factor of ten or less, as compared to
+the original CPU implementation.
+However, we demonstrated that additional orders-of-magnitude
+improvements could be achieved by removing that limitation.
+We also discussed a conceivable solution to cope with that limitation,
+using a concept of the Imprecise Computation model.
+
+In future work, we extend our scheme to multidimensional OV models.
+This extention allows our research to be applied to a broader range of
+problems beyond freeways traffic simulation.
+We also plan to consider a specific approach to the integration of the
+Imprecise Computation model into our scheme in order to facilitate
+deployment of traffic simulation in the real world.
+Another interesting direction in this line of work is to coordinate with
+engineering-based simulation such as AutoMatrix~\cite{Mangharam2011}.
 

### 2012-09-16 draft.tex
changed the title
refactored the paper

@@
 
-\title{Accelerating Traffic Simulation and Its Real-Time Challenges}
+\title{Toward GPU-accelerated Traffic Simulation and Its Real-Time Challenge}
 
 \author {
-Manato Hirabayashi, Shinpei Kato and Masato Edahiro\\
+Manato Hirabayashi, Shinpei Kato, Masato Edahiro\\
 \textit{Department of Information Engineering}\\
@@
 Yuki Sugiyama\\
-\textit{Department of Physics}\\
+\textit{Department of Complex Systems Science}\\
 \textit{Nagoya University}\\
@@
  Traffic simulation is a growing domain of computational physics.
- There are many applications in our industry and life that benefit from
- traffic simulation.
- A core challenge of this research domain, however, is posed due to
+ There are many applications in our life and industry that benefit
+ from traffic simulation.
+ A core challenge of this science research, however, is posed due to
  an unbounded scale of computation.
@@
  Two schemes are presented to study how to maximize GPU
- performance in this context.
- Our evaluation using a commodity graphics card shows that our GPU
- implementation improves simulation time by about five times over the
- original CPU implementation.
- We also demonstrate by experiments that additional orders-of-magnitude
+ performance in the context of traffic simulation.
+ Our experimental results show that even the most basic GPU
+ implementation based on our schemes accelerates simulation by about
+ five times over the original CPU implementation.
+ We also demonstrate that additional orders-of-magnitude
  improvements could be achieved by overcoming the current hardware
@@
 
-Transportation systems underlie our industry and life as part of
+Transportation systems underlie our life and industry, as part of
 societal infrastructure.
-Economy depends highly on traffic flow.
-Once a traffic jam occurs, for example, it could cause significant
-economic damage through degradation of transport efficiency, energy
-consumption, and environmental poisoning.
-According to the Japanese government report a few years ago, economic
-losses due to traffic congestion reach one hundred billion dollars per
-year in Japan.
+Traffic flow is particularly a major factor that influences economy.
+According to the Japanese government report, economic losses due to
+traffic congestion could reach one hundred billion dollars per year in
+Japan, through degradation of transport efficiency, energy consumption,
+and environmental poisoning.
 Albeit a major issue of transportation systems, the mechanism of traffic
-jam is not well-explained in the literature.
+congestion is not well-explained in the literature.
 What is particularly important is ``phantom jam'', which often happens
@@
 a physical phenomenon but not one caused by human errors.
-However, we are not aware of scientific and engineering solutions to
-such real-world problems.
+However, we have not been aware of scientific and engineering solutions
+to such real-world problems yet.
 
 In physics, traffic flow is described by mathematical
-models~\cite{Bando1995, Kerner1993, Nagel1992}.
-While their mathematical expressions are not identical, they are all
-compute-intensive procedures.
+models~\cite{Bando1995, Bando1995_2, Kerner1993, Nagel1992}.
+While their mathematical expressions are not identical, they are
+commonly compute-intensive forms.
 For example, traffic simulation based on the Optimal Velocity (OV)
-model~\cite{Bando1995} computes locations and velocities of agents every
-sampling period, solving OV equations. 
+model~\cite{Bando1995, Bando1995_2} computes locations and velocities of
+agents every sampling period, solving OV equations. 
 Specifically, the location $x_n$ of the $n$th agent is described by the
@@
 
-Deployment of traffic simulation may require real-time feedback from the
-real world.
-This turns out to be a cyber-physical systems (CPS) problem.
+Computational workload is not the only issue of concern.
+Deployment of traffic simulation may also require real-time feedback
+from the real world, which turns out to be a cyber-physical systems
+(CPS) problem indeed.
 Real-time traffic simulation is another challenging issue, where
@@
 requirement of a given scenario.
-For instance, evacuation may want very high rate simulation, sacrificing
-the preciseness of simulation to some extent.
+For instance, simulation of emergent evacuation may want very high-rate
+computation, even sacrificing preciseness of simulation to some extent.
 Unfortunately, none of those challenges has been explored yet, largely
-due to a lack of multidisciplinary research collaborations between physics
-and computer science.
+due to a lack of multidisciplinary collaborations between computationsl
+physics and computer science.
 
@@
 The rest of this paper is organized as follows.
-Section~\ref{sec:assumption} describes the assumption and terminology
-used this paper.
-Section~\ref{sec:traffic_simulation} briefly explains our traffic
-simulation program.
+Section~\ref{sec:assumption} describes the basic assumption and
+terminology used in this paper.
+Section~\ref{sec:traffic_simulation} explains an overview of
+traffic simulation and its literature.
 Section~\ref{sec:gpu_implementations} presents our schemes of GPU
-implementations of traffic simulation.
-Section~\ref{sec:evaluation} evaluates the benefit of our GPU
-implementations, and provides an insight into future work.
+implementations for traffic simulation.
+Section~\ref{sec:evaluation} evaluates the advantage of using our
+schemes, and provides an insight into future work.
 Section~\ref{sec:conclusion} concludes this paper.
@@
 
+While we are intensively interested in CUDA and GPUs, the notion of
+accelerating traffic simulation presented in this paper is widely
+applicable to heterogeneous architectures, where CPUs and 
+accelerators co-exist to complement each other. 
+GPUs are currently the most well-recognized forms of accelerators, but
+emerging alternatives include the Intel Many Integrated Core (MIC) and
+AMD Fusion architectures.
+However, their programming models are almost identical in that a master
+thread running on the CPU manages the program flow, while parallelized
+threads are offloaded on to accelerators to boost computing performance.
+CUDA, OpenCL, HMPP, and AMP are all such programming languages that
+support heterogeneous architectures under this concept.
+
 \section{Traffic Simulation}
@@
 physical models exist.
-This paper is particularly focused on the OV model~\cite{Bando1995}.
+This paper is particularly focused on the OV model~\cite{Bando1995,
+Bando1995_2}.
 In this model, a driver is supposed to maintain an optimal velocity that
@@
 To begin with, however, we restrict our attention to the OV model.
-More details are found in~\cite{Bando1995}.
+More details are found in~\cite{Bando1995, Bando1995_2}.
 
 The traffic simulation program based on the OV model is straightforward,
-as the model itself has a simple form:
+as the model itself has a simple form.
+In this paper, we abstract its algorithm by the following stages:
 
@@
 
-The computation in Step 3 exploits a lot of loop procedures to derive
-locations and velocities of agents.
-This part is the most compute-intensive, and thus can be accelerated by
-the GPU.
+The computational workload in Step 3 exploits a lot of loop executions
+to derive locations and velocities of agents.
+This is the most compute-intensive part of the algorithm.
+The primary goal of this paper is to find out the parallelism of this
+workload that can be applied to the GPU.
+By nature, traffic simulation based on the OV model is very scalable in
+terms of performance with respect to the number of agents.
+The location and velocity of each agent depends on only those of
+neighbors.
+It is worth exploring a breakthrough of this research domain by
+coordinating computational physics and computer science technology.
 
@@
 
-GPU performance is dominated by the program design.
+The primary contribution of this paper is to identify the fundamental
+schemes of GPU implementations for traffic simulation.
+It is very important to understand that GPU performance is dominated by
+the program design.
+We first reason about the motivation of our schemes.
+
 GPU-accelerated programs are typically divided into two pieces of code.
@@
 
-We focus on the one-dimensional OV model~\cite{Bando1995}, which can be
-applied to freeways simulation.
+We focus on the one-dimensional OV model~\cite{Bando1995, Bando1995_2},
+which can be applied to freeways simulation.
 The original program~\cite{OV_Program} of this simulation is written in
@@
 requirement for users.
-Now, this turns into a real-time systems problem.
-We have developed a lot of solutions to use the GPU in real-time
+Imagine how many potential users exist in the real-world transportation
+system.
+We are facing a core challenge of next-generation real-time systems.
+
+We have developed basic solutions to use the GPU in real-time
 systems~\cite{Kato2011_3, Kato2011_2, Kato2011_1, Kato2012}.
-We believe that some of these work could contribute to GPU-accelerated
-real-time traffic simulation.
+Resource management with scheduling, synchronization, and memory
+management of GPU applications has been addressed in our previous
+work.
+We believe that a grander vision of GPU-accelerated traffic
+simulation will need an integration of cutting-edge real-time systems
+technology and computational physics.
 

### 2012-09-16 draft.tex
minor fixes

@@
 Unfortunately, none of those challenges has been explored yet, largely
-due to a lack of multidisciplinary collaborations between computationsl
+due to a lack of multidisciplinary collaborations between computational
 physics and computer science.
@@
 $\Delta t$.
-The value of $\Delta t$ can be determined adaptively based on the
+The value of $\Delta t$ can be determined dynamically based on the
 Imprecise Computation model.
@@
 In future work, we extend our scheme to multidimensional OV models.
-This extention allows our research to be applied to a broader range of
+This extension allows our research to be applied to a broader range of
 problems beyond freeways traffic simulation.

### 2012-11-04 draft.tex
futher minor fix
updated for the camera-ready

@@
  Traffic simulation is a growing domain of computational physics.
- There are many applications in our life and industry that benefit
- from traffic simulation.
- A core challenge of this science research, however, is posed due to
- an unbounded scale of computation.
- In this paper, we explore an advantage of using the graphics processing
- unit (GPU) to support its computation.
- Two schemes are presented to study how to maximize GPU
- performance in the context of traffic simulation.
- Our experimental results show that even the most basic GPU
- implementation based on our schemes accelerates simulation by about
- five times over the original CPU implementation.
- We also demonstrate that additional orders-of-magnitude
- improvements could be achieved by overcoming the current hardware
- limitation.
+ Many life and industrial applications would benefit from traffic
+ simulation to establish reliable transportation systems.
+ A core challenge of this science research, however, is its unbounded
+ scale of computation.
+ This paper explores an advantage of using the graphics processing
+ unit (GPU) for this computational challenge.
+ We study two schemes of maximizing GPU performance in the context of
+ traffic simulation, and provide some basic experiments.
+ The experimental results show that our GPU implementation improves
+ simulation speed by five times over the traditional CPU
+ implementation.
+ We also discuss that additional orders-of-magnitude improvements could
+ be achieved by overcoming the current hardware limitation of the GPU.
 \end{abstract}
@@
 
-Transportation systems underlie our life and industry, as part of
-societal infrastructure.
-Traffic flow is particularly a major factor that influences economy.
-According to the Japanese government report, economic losses due to
-traffic congestion could reach one hundred billion dollars per year in
-Japan, through degradation of transport efficiency, energy consumption,
-and environmental poisoning.
+Transportation systems are part of important infrastructures for our life
+and industry.
+For example, traffic flow highly influences economy.
+According to the report from the Japanese government, economic losses
+due to traffic congestion could reach one hundred billion dollars per
+year in Japan, which include degradation of transport efficiency, energy
+consumption, and environmental poisoning.
 Albeit a major issue of transportation systems, the mechanism of traffic
 congestion is not well-explained in the literature.
-What is particularly important is ``phantom jam'', which often happens
-on freeways without any accident.
-This phantom jam should be avoided by science, given that it is
-a physical phenomenon but not one caused by human errors.
-However, we have not been aware of scientific and engineering solutions
-to such real-world problems yet.
+Examples include a ``phantom jam'' -- a traffic jam that naturally
+occurs on freeways without any car and/or human accidents.
+This phantom jam should be addressed by science, since it is a physical
+phenomenon but not one caused by human errors.
+There is certainly a challenge of real-time simulation in this line of
+work, but even a basic scale of execution times is not understood at all.
 
@@
 it is apparent that computational workload increases exponentially.
-In fact, the above formula considers only one dimension, which restricts
+The above formula considers only one dimension, which restricts
 applications of simulation to freeway traffic flows, powder flows,
@@
 animals, to use similar optimal velocity models.
-Given a scale of million agents in the real world, traffic
-simulation should be supported by powerful computer systems.
+Given a scale of million agents in the real world, traffic simulation
+must be supported by powerful computer systems.
 
@@
 This paper explores how to accelerate traffic simulation using the
-state-of-the-art parallel computing technology.
-In particular, we use the graphics processing unit (GPU), which
-integrates hundreds of cores on a chip.
+graphics processing unit (GPU), which integrates hundreds of processing
+cores on a chip.
 Recent GPUs are becoming more and more suitable for general-purpose
@@
 The resulting workload is highly data-parallel and compute-intensive,
-which can be nicely offloaded on to the GPU.
-We also identify the bottleneck in accelerating our traffic simulation
-program using the GPU, and provide some insight into its solution.
+which can be offloaded on to the GPU.
+We also identify a bottleneck of traffic simulation when accelerated by
+the GPU, and provide an insight into some solutions.
+The main objective of this paper is to figure out a computational scale
+of traffic simulation.
+Timing guarantees are not considered, but our contribution is essential
+to achieve real-time traffic simulation in future work.
 
@@
 Section~\ref{sec:assumption} describes the basic assumption and
-terminology used in this paper.
+terminology behind this paper.
 Section~\ref{sec:traffic_simulation} explains an overview of
@@
 
-While we are intensively interested in CUDA and GPUs, the notion of
-accelerating traffic simulation presented in this paper is widely
-applicable to heterogeneous architectures, where CPUs and 
-accelerators co-exist to complement each other. 
+This paper is particularly focused on CUDA and GPUs, but the notion of
+accelerating traffic simulation presented in this paper is applicable to
+heterogeneous architectures that integrate CPUs and accelerators.
 GPUs are currently the most well-recognized forms of accelerators, but
@@
 thread running on the CPU manages the program flow, while parallelized
-threads are offloaded on to accelerators to boost computing performance.
-CUDA, OpenCL, HMPP, and AMP are all such programming languages that
-support heterogeneous architectures under this concept.
+threads are offloaded on to accelerators.
+CUDA, OpenCL, HMPP, and AMP are all based on this programming model.
 
@@
 system.
-We are facing a core challenge of next-generation real-time systems.
+We are now facing a core challenge of next-generation real-time
+systems.
 
-We have developed basic solutions to use the GPU in real-time
-systems~\cite{Kato2011_3, Kato2011_2, Kato2011_1, Kato2012}.
-Resource management with scheduling, synchronization, and memory
-management of GPU applications has been addressed in our previous
+Recently real-time systems have been augmented with GPU resource
+management techniques~\cite{Elliott_RTS12, Elliott_ECRTS12, Kato2011_3,
+Kato2011_2, Kato2011_1, Kato2012, Liu_PACT12}.
+Scheduling, synchronization, and memory management of GPU applications
+are becoming key challenges to applying the GPU for real-time systems.
+A grander vision of GPU-accelerated traffic simulation will need a
+tighter integration of real-time systems technology and computational
+physics.
+We believe that this paper plays a vital role to facilitate this line of
 work.
-We believe that a grander vision of GPU-accelerated traffic
-simulation will need an integration of cutting-edge real-time systems
-technology and computational physics.
 
@@
 deployment of traffic simulation in the real world.
-Another interesting direction in this line of work is to coordinate with
+This would pose a core challenge of real-time systems.
+Another interesting direction of future work is a coordination with
 engineering-based simulation such as AutoMatrix~\cite{Mangharam2011}.
+
+\section*{Acknowledgement}
+
+This work was supported by Grant-in-Aid for Scientific Research
+24500058.
 
